Towards a standardized evaluation of imputation methodology

### 2022-03-23 Formatted/Formatted.tex
Merge duplicate info about data generation from problems and solutions sections

To evaluate the ability of an imputation routine to handle missingness, a form of truth has to be established. Those who perform simulation studies are in the luxury position to establish the truth beforehand through a data generating mechanism which yields complete data. The data generating mechanism is typically either in line with model-based or design-based simulation. With the former, data are generatd from a statistical model. With the latter, data are sampled from a sufficiently large observed set.

The problem with model-based data generating mechanisms is that the performance of imputation methods may not translate to empirical data. With this approach, the complete data are drawn from a known (parametric) model, whereas real-life data hardly ever follow a given theoretical distribution. The generalizibility of performance outside of the simulation setting is at an even greater risk when data are generated following a model that is also used for imputing the data. The performance of this specific imputation approach is then deemed good (or better than other methods), purely due to the evaluated conditions being in favor of the problem that is studied. ~~Although such procedures may be statistically relevant, the approach would be no good for even the simplest empirical case.~~ Moreover, data are often generated such that the problem that is being studied is most pronounced. This results in simulated data that contain very valuable information structures, e.g. consistently high correlations between groups of variables. Then, no matter what type of missingness is induced, the observed parts of the data still hold much, if not all, of the information about the missing part and the performance of the imputation procedure(s) is not surprisingly evaluated as very good. ~~Also, it may be that the relations in real-life data are rarely as pronounced as the problem that is being studied.~~ [maybe end in something positive? or some other bridge to the next paragraph?]

@@

% For evaluations of model-based simulations, it could be convincing to demonstrate a method's applicability to real-world missing data problems. This can, for example, be achieved by obtaining and imputing a fully observed set of variables, wherein the missingness is mimicked from a similar, incomplete set. Alternatively, an incomplete set could be obtained and truth could be established by removing the incomplete cases from the data. However, the real-world missingness would then be omitted.

@@

% Irrespective of the data generating mechanism,

Problems with either of these data generation approaches is that the number of observations, the number of variables, the variable type(s) and the coherence between variables may give some imputation methods a comparative advantage, if they are not a) varied across simulation conditions [when is this fair? is this even an option?], or b) the object of the study [add some examples of simulations specifically focused on small samples, categorical data, etc.].

Something else to consider is that if missing data methodology is the primary focus of the simulation, it may not be necessary to draw data repeatedly (whether it may be from a known model or empirical data). [Add pooling when the sample is the population citep{vink14} and write something about bootstrapping SEs and that SEs may be incorrect if the (re)sampling is not handled correctly.]

For situations where sampling variance is not of interest, the sampling process can be omitted and a single complete dataset can simply be obtained by model- or design-based approaches. The parameters of that single complete set will serve as comparative truth. This process is computationally convenient, because only a single complete datasets has to be considered during all of the simulations. One needs to realize, however, that the conventional pooling rules \citep[cf.][p. 76-77]{rubi87} do not apply for finite population inference and that alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}. [in our own experience it may be difficult to get true parameter estimates of transformed data (e.g., the mean of a variable with skewness).]

@@

[In geveral, adhere to ADEMP \citep{morr18}. Make aim clear. Choose scope (type of DGM, target) and imputation methods to evaluate. Ideally, perform fully factorial design, see \citet{morr18}. Be explicit in choices, better even to share the script directly.]

@@

\subsection{Step 1. Obtain truth}

@@

The simulator must to decide whether there is a need for sampling variance in the simulations scheme. If sampling variance cannot be omitted from the simulation scheme, two approaches are possible:
\begin{enumerate}
  \item Model-based simulation: draw samples from a known probability distribution, such as the multivariate normal distribution. The theoretical parameters under which the samples are obtained will serve as the comparative truth during the simulations. 
  \item Design-based simulation: sample data from a sufficiently large set. The parameters of the sufficiently large set will serve as the comparative truth during the simulations. This approach is often used in situations where a probability distribution is not available, or where real-life data structures are of interest. Applications of design-based simulation are often found in official statistics. A benefit of design-based simulation is the ability to use real-life observed data structures. 
\end{enumerate}
For situations where sampling variance is not of interest, the sampling process can be omitted and a single complete dataset can simply be obtained by model- or design-based approaches. The parameters of that single complete set will serve as comparative truth. This process is computationally convenient, because only a single complete datasets has to be considered during all of the simulations. One needs to realize, however, that the conventional pooling rules \citep[cf.][p. 76-77]{rubi87} do not apply for finite population inference and that alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}.

% we hebben al sampling variance tussen de incomplete datasets, extra sampling variance op het complete data level vertroebelt het proces alleen maar. we verplaatsen liever de monte carlo naar het niveau van de missingness.

@@

\subsection{Step 2. Induce missingness}

@@

\subsection{Step 3. Apply missing data method(s) and analysis model(s)}

@@

Report the following.

[Put this in a table, like this one]

\begin{table}[htb]
\begin{center}
\caption{The caption of a table.}
\begin{tabular}{lll}
\hline
Description 1 & Description 2 & Description 3\\
\hline
Row 1, Col 1 & Row 1, Col 2 & Row 1, Col 3\\
Row 2, Col 1 & Row 2, Col 2 & Row 2, Col 3\\
\hline
\end{tabular}
\end{center}
\end{table}

@@

>>>>

To evaluate the ability of an imputation routine to handle missingness, a form of truth has to be established. Those who perform simulation studies are in the luxury position to establish the truth beforehand by choosing a data generating mechanism. Data generating mechanisms define how a complete dataset is obtained at the start of each simulation repetition. There are two general approaches to generaring complete data: (i) model-based simulation, in which data are drawn from a known statistical model or probability distribution, such as the multivariate normal distribution; and (ii) design-based simulation, where data are sampled (with replacement) from a sufficiently large observed set, such as from official statistics records.

The problem with model-based data generating mechanisms is that a method's performance on simulated data may not translate to empirical data. Real-life data hardly ever follow a given theoretical distribution, so there is no guarantee that simulation results are generalizable. Moreover, data are often generated such that the problem that is being studied is most pronounced, e.g. with consistently high correlations between groups of variables. This results in simulated data that contain such valuable information structures that, no matter what type of missingness is subsequently induced, the observed parts of the data still hold much (if not all) of the information about the missing part. Unsurprisingly, the performance of any imputation method will then be evaluated as good. Another threat to the generalizibility of model-based simulations is the use of a single model for both data generation and imputation. If data are generated following a model that is also used for imputing the data, the imputation approach will be deemed good (or better than other methods) purely due to the evaluated conditions being in favor of the problem that is studied. Other (unfair) comparative advantages in favor of a certain imputation method may occur due to characteristics of the generated data, such as the number of observations, the number of variables, the variable type(s) and the coherence between variables. In contrast to design-based studies, such characteristics are not always explicit conditions of the data, which may give a false sense of objectivity.

@@

@@

Something else to consider is that if missing data methodology is the primary focus of the simulation, it may not be necessary to draw data repeatedly (whether it may be from a known model or empirical data). [Add pooling when the sample is the population citep{vink14} and write something about bootstrapping SEs and that SEs may be incorrect if the (re)sampling is not handled correctly.]  One needs to realize that the conventional pooling rules \citep[cf.][p. 76-77]{rubi87} do not apply for finite population inference and that alternative pooling rules need to be used \citep{raghunathan2003multiple,vink14}. [in our own experience it may be difficult to get true parameter estimates of transformed data (e.g., the mean of a variable with skewness).] % we hebben al sampling variance tussen de incomplete datasets, extra sampling variance op het complete data level vertroebelt het proces alleen maar. we verplaatsen liever de monte carlo naar het niveau van de missingness.

@@

[In geveral, adhere to ADEMP \citep{morr18}. But there are some differences for the specific case of imputation methodology. All of the parts in ADEMP are used, but a bit scrambled up. Table 1 outlines some of the steps to consider in simulation studies focused on evaluating imputation methodology. The rest of this section is ordered based on these steps.]

\begin{table}[htb]
\begin{center}
\caption{Steps to consider in imputation simulation studies.}
\begin{tabular}{lll}
\hline
Step                  & Aspects under consideration \\
\hline
1. Set scope          & Aim(s), missing data method(s) to evaluate, number of simulation \\
                      & repetitions \\
2. Obtain truth       & Data generating mechanism(s), estimand(s), sampling variance \\
3. Induce missingness & Missingness mechanism(s), missing data pattern(s), missingness \\
                      & proportion(s) \\
4. Apply methods      & Imputation model(s), analysis model(s) \\
% Benchmarking, imputation model parameter, estimand
5. Evaluate methods   & Imputation model fit, performance measures \\
6. Report             & Text, visualization, checklist, simulation script \\
\hline
\end{tabular}
\end{center}
% \footnotesize{Note. Multiple aspects may be varied across simulation conditions (e.g., using a factorial design with several data generating mechanisms and missingness proportions).}
\end{table}

@@

\subsection{Set scope}

Make aim clear. Choose scope (type of DGM, target) and imputation methods to evaluate. Ideally, perform fully factorial design, see \citet{morr18}. Think about tolerable level of uncertainty to pick a number of simulation repetitions. Think about general set-up: are the missing data methods all applied to the same incomplete dataset in each simulation repetition, or is a new incomplete set created per method?

\subsection{Obtain truth}

@@

[Decide which data genetation mechanism to use and obtain true value of the estimand(s).] In model-based simulation data are drawn from a known probability distribution, so the theoretical parameters under which the samples are obtained serve as the comparative truth in the simulations. In design-based simulation, data are sampled from a sufficiently large set, so the parameters of this specific set serve as the comparative truth in the simulations.

The design-based approach is often used in situations where a probability distribution is not available, or where real-life data structures are of interest. Applications of design-based simulation are often found in official statistics. A benefit of design-based simulation is the ability to use real-life observed data structures.

The simulator must to decide whether there is a need for sampling variance in the simulations scheme. For situations where sampling variance is not of interest, the sampling process can be omitted and a single complete dataset can simply be obtained from the data generating mechanism. The parameters of that single complete set will serve as comparative truth. This process is computationally convenient, because only a single complete datasets has to be considered during all of the simulations. Use alternative pooling rules for finite population inference \citep{raghunathan2003multiple,vink14}. If sampling variance cannot be omitted from the simulation scheme, draw data repeatedly and apply conventional pooling rules \citep[cf.][p. 76-77]{rubi87}.

@@

\subsection{Induce missingness}

@@

\subsection{Apply missing data method(s) and analysis model(s)}

@@

Be explicit in choices, better even to share the script directly.

Describe the simulation textually, use pseudo-code, a flowchart, visualize missingness and simulation results.

Fill in the following checklist.

@@

This document is aimed at establishing a common ground for the evaluation of imputation routines. Such a common ground would be the basis of a standardized evaluation. This allows for fair and efficient comparisons between imputation techniques. Ultimately, it would be desirable to evaluate every imputation routine against the same standardized set in order to quantify the statistical properties across imputation routines. If properly executed, this would allow for careful matching of imputation methodologies to new missing data problems.
